---
title: "16S Alpha Diversity Analysis"
author: "Jacob T. Nearing"
date: "2/15/2021"
output: 
  rmdformats::downcute:
    self_contained: true
    default_style: "light"
    downcute_theme: "chaos"
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(lmerTest)
library(ggplot2)
library(emmeans)
library(dplyr)
library(broom)
library(kableExtra)
library(cowplot)
library(reshape2)
library(dunn.test)
library(ggpubr)
library(ggbeeswarm)
library(FSA)
colors <- c('ferrous bisglycinate'="#a09342", 'ferrous sulfate'="#9673c5", 'placebo'="#ba6d75")
```

# Introduction

In this analysis we will examine the relationship between four different alpha diversity metrics and the use of iron supplementation in this cohort.

We will examine:

Richness: Examiens the total number of "unique" members identified.
Evenness: How equally spread the abundance is across different members of the community
Shannon diversity: Looks at both richness and evenness in a single metric
Faith's phylogenetic diversity: Measures how phylogenetically related the community members are while also taking into their abundance

```{r}
Metadata <- read.table("~/projects/IRON_SUP_EMMA/Data/Metadata/Iron_Metadata_21_03_09.txt", header=T, sep="\t", row.names=1)

## PD

Faith_pd <- read.table("~/projects/IRON_SUP_EMMA/Data/diversity_5000/raw_alpha/faith_pd/alpha-diversity.tsv",
                       row.names = 1, header=T)

## evenness
Evenness <- read.table("~/projects/IRON_SUP_EMMA/Data/diversity_5000/raw_alpha/evenness/alpha-diversity.tsv",
                       row.names = 1, header=T)


## shannon
Shannon <- read.table("~/projects/IRON_SUP_EMMA/Data/diversity_5000/raw_alpha/shannon/alpha-diversity.tsv",
                      row.names = 1, header=T)

## raw_asvs
Raw_ASVs <- read.table("~/projects/IRON_SUP_EMMA/Data/diversity_5000/raw_alpha/raw_ASVs/alpha-diversity.tsv",
                       row.names = 1, header=T)


Alpha_data <- cbind(Faith_pd, Evenness, Shannon, Raw_ASVs)

Metadata_filter <- Metadata[rownames(Alpha_data),]
```

# Change in diversity metric {.tabset}
## Faith's Dive

```{r}
comparisons <- list(c("ferrous bisglycinate\nn=40", "ferrous sulfate\nn=46"),
                    c("ferrous bisglycinate\nn=40", "placebo\nn=47"),
                    c("ferrous sulfate\nn=46", "placebo\nn=47"))
comb_data <- cbind(Alpha_data, Metadata_filter)
## we need to remove samples that do not have a match up... 

faith_melted_metadata_filt <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("faith_pd"))
faith_melted_metadata_filt_filt <- faith_melted_metadata_filt %>% group_by(Patint_ID) %>% filter(n()==2) %>% mutate(slope=value[Week==12] - value[Week==0])

#remove all week12
faith_melted_metadat_pairwise <- faith_melted_metadata_filt_filt[-which(faith_melted_metadata_filt_filt$Week==12),]

#faiths
boxplot(faith_melted_metadat_pairwise$slope ~ faith_melted_metadat_pairwise$arm, xlab="Arm of Treatment", ylab="Change in Faith's PD")

## make a nice ggplot for pub

sample_size = faith_melted_metadat_pairwise %>% group_by(arm) %>% summarize(n())
sample_size


faith_melted_metadat_pairwise <- faith_melted_metadat_pairwise %>% 
  left_join(sample_size) %>% 
  left_join(data.frame(colors, arm=names(colors))) %>%
  mutate(axis_name = paste0(arm, "\n", "n=", `n()`))

faith_change_simple <- faith_melted_metadat_pairwise %>%
  ggplot(aes(x=axis_name, y=slope)) +
  geom_violin(alpha=0.3, aes(fill=colors)) +
  geom_boxplot(width=0.1, fill="white", outlier.shape = NA) +
  scale_fill_identity() +
  xlab("") +
  ylab("Change in Faith's PD") +
  #stat_compare_means(label.y=15, hide.ns=T) +
  geom_quasirandom() +
  theme_cowplot()

faith_change_simple
```


Run pairwise wilcox test using placebo as reference group.

### add manual dunn's stat results!

```{r}
faith_pairwise_res <- compare_means(slope ~ axis_name, data=faith_melted_metadat_pairwise,
              p.adjust.method = "holm", method="wilcox.test", ref.group = "placebo\nn=47") %>% mutate(y.position=c(10,11))


faith_change_simple + stat_pvalue_manual(faith_pairwise_res)

### not significant!
```


## Shannon Diversity
```{r}
shannon_melted_metadata_filt <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("shannon_entropy"))
shannon_melted_metadata_filt_filt <- shannon_melted_metadata_filt %>% group_by(Patint_ID) %>% filter(n()==2) %>% mutate(slope=value[Week==12] - value[Week==0])

#remove all week12
shannon_melted_metadat_pairwise <- shannon_melted_metadata_filt_filt[-which(shannon_melted_metadata_filt_filt$Week==12),]

#faiths
boxplot(shannon_melted_metadat_pairwise$slope ~ shannon_melted_metadat_pairwise$arm, xlab="Arm", ylab="Change in Shannon Diversity")

shannon_melted_metadat_pairwise <- shannon_melted_metadat_pairwise %>% 
  left_join(sample_size) %>% 
  left_join(data.frame(colors, arm=names(colors))) %>%
  mutate(axis_name = paste0(arm, "\n", "n=", `n()`))

shannon_change_simple <- shannon_melted_metadat_pairwise %>% 
  ggplot(aes(x=axis_name, y=slope)) +
  geom_violin(alpha=0.3, aes(fill=colors)) +
  geom_boxplot(width=0.1, fill="white", outlier.shape = NA) +
  scale_fill_identity() +
  xlab("") +
  ylab("Change in Shannon Diversity") +
  #stat_compare_means(label.y=7) +
  geom_quasirandom() +
  theme_cowplot()

shannon_change_simple
```

```{r}
shannon_pairwise_res <- compare_means(slope ~ axis_name, data=shannon_melted_metadat_pairwise,
              p.adjust.method = "holm", method="wilcox.test", ref.group = "placebo\nn=47") %>% mutate(y.position=c(4,4.5))


shannon_change_simple + stat_pvalue_manual(shannon_pairwise_res)

```


## Evenness
```{r}
even_melted_metadata_filt <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("pielou_evenness"))
even_melted_metadata_filt_filt <- even_melted_metadata_filt %>% group_by(Patint_ID) %>% filter(n()==2) %>% mutate(slope=value[Week==12] - value[Week==0])

#remove all week12
even_melted_metadat_pairwise <- even_melted_metadata_filt_filt[-which(even_melted_metadata_filt_filt$Week==12),]

#faiths
boxplot(even_melted_metadat_pairwise$slope ~ even_melted_metadat_pairwise$arm, xlab="Arm", ylab="Change in Evenness")

even_melted_metadat_pairwise <- even_melted_metadat_pairwise %>% 
  left_join(sample_size) %>% 
  left_join(data.frame(colors, arm=names(colors))) %>%
  mutate(axis_name = paste0(arm, "\n", "n=", `n()`))

even_change_simple <- even_melted_metadat_pairwise %>% 
  ggplot(aes(x=axis_name, y=slope)) +
  geom_violin(alpha=0.3, aes(fill=colors)) +
  geom_boxplot(width=0.1, fill="white", outlier.shape = NA) +
  scale_fill_identity() +
  xlab("") +
  ylab("Change in Evenness") +
  #stat_compare_means(label.y=.75) +d
  geom_quasirandom() + 
  theme_cowplot()

even_change_simple
```

```{r}
even_pairwise_res <- compare_means(slope ~ axis_name, data=even_melted_metadat_pairwise,
              p.adjust.method = "holm", method="wilcox.test", ref.group = "placebo\nn=47") %>% mutate(y.position=c(.7,.8))


even_change_simple + stat_pvalue_manual(even_pairwise_res)
## none are significant
```


## Richness
```{r}
rich_melted_metadata_filt <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("observed_features"))
rich_melted_metadata_filt_filt <- rich_melted_metadata_filt %>% group_by(Patint_ID) %>% filter(n()==2) %>% mutate(slope=value[Week==12] - value[Week==0])

#remove all week12
rich_melted_metadat_pairwise <- rich_melted_metadata_filt_filt[-which(rich_melted_metadata_filt_filt$Week==12),]

#faiths
boxplot(rich_melted_metadat_pairwise$slope ~ rich_melted_metadat_pairwise$arm, xlab="Arm", ylab="Change in Richness")


rich_melted_metadat_pairwise <- rich_melted_metadat_pairwise %>% 
  left_join(sample_size) %>% 
  left_join(data.frame(colors, arm=names(colors))) %>%
  mutate(axis_name = paste0(arm, "\n", "n=", `n()`))

rich_change_simple <- rich_melted_metadat_pairwise %>% 
  ggplot(aes(x=axis_name, y=slope)) +
  geom_violin(alpha=0.3, aes(fill=colors)) +
  geom_boxplot(width=0.1, fill="white", outlier.shape = NA) +
  scale_fill_identity() +
  xlab("") +
  ylab("Change in Richness") +
  #stat_compare_means(label.y=250) +
  geom_quasirandom() +
  theme_cowplot()

rich_change_simple
```

```{r}
rich_pairwise_res <- compare_means(slope ~ axis_name, data=rich_melted_metadat_pairwise,
              p.adjust.method = "holm", method="wilcox.test", ref.group = "placebo\nn=47") %>% mutate(y.position=c(165,185))


rich_change_simple + stat_pvalue_manual(rich_pairwise_res)
```

## Publication figure

```{r}
Figure1_top_no_stats <- plot_grid(rich_change_simple, even_change_simple, nrow=1)
Figure1_top_no_stats
saveRDS(Figure1_top_no_stats, "~/GitHub_Repos/Cambodia_Iron_Supplement/Display_items/Intermediate_RDS_files/Figure1_top_no_stats.RDS")



Figure1_top_yes_stats <- plot_grid(rich_change_simple + stat_pvalue_manual(rich_pairwise_res),
                                   even_change_simple + stat_pvalue_manual(even_pairwise_res),
                                   nrow=1)
Figure1_top_yes_stats
saveRDS(Figure1_top_yes_stats, "~/GitHub_Repos/Cambodia_Iron_Supplement/Display_items/Intermediate_RDS_files/Figure1_top_yes_stats.RDS")


Supplemental_Fig1_no_stats <- plot_grid(rich_change_simple, even_change_simple, faith_change_simple, shannon_change_simple)

ggsave(filename = "~/GitHub_Repos/Cambodia_Iron_Supplement/Display_items/Supp_Figures/Supplemental_Figure1_no_stats.pdf",
       plot=Supplemental_Fig1,
       width=9,
       height=10,
       dpi=600)


Supplemental_Fig1_yes_stats <- plot_grid(rich_change_simple + stat_pvalue_manual(rich_pairwise_res),
                                         even_change_simple + stat_pvalue_manual(even_pairwise_res),
                                         faith_change_simple + stat_pvalue_manual(faith_pairwise_res),
                                         shannon_change_simple + stat_pvalue_manual(shannon_pairwise_res))

ggsave(filename = "~/GitHub_Repos/Cambodia_Iron_Supplement/Display_items/Supp_Figures/Supplemental_Figure1_yes_stats.pdf",
       plot=Supplemental_Fig1_yes_stats,
       width=9,
       height=10,
       dpi=600)
```

# Examination of complient individuals only

```{r}
Metadata_filter_comp <- Metadata_filter[which(Metadata_filter$Compliant==1),]
Alpha_data_comp <- Alpha_data[rownames(Metadata_filter_comp),]

```

## Faith's Phylogenetic Diversity
```{r}
comb_data <- cbind(Alpha_data_comp, Metadata_filter_comp)
## we need to remove samples that do not have a match up... 

faith_melted_metadata_filt <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("faith_pd"))
faith_melted_metadata_filt_filt <- faith_melted_metadata_filt %>% group_by(Patint_ID) %>% filter(n()==2) %>% mutate(slope=value[Week==12] - value[Week==0])

#remove all week12
faith_melted_metadat_pairwise <- faith_melted_metadata_filt_filt[-which(faith_melted_metadata_filt_filt$Week==12),]

#faiths
boxplot(faith_melted_metadat_pairwise$slope ~ faith_melted_metadat_pairwise$arm, xlab="Arm of Treatment", ylab="Change in Faith's PD")
```

## Shannon Diversity
```{r}
shannon_melted_metadata_filt <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("shannon_entropy"))
shannon_melted_metadata_filt_filt <- shannon_melted_metadata_filt %>% group_by(Patint_ID) %>% filter(n()==2) %>% mutate(slope=value[Week==12] - value[Week==0])

#remove all week12
shannon_melted_metadat_pairwise <- shannon_melted_metadata_filt_filt[-which(shannon_melted_metadata_filt_filt$Week==12),]

#faiths
boxplot(shannon_melted_metadat_pairwise$slope ~ shannon_melted_metadat_pairwise$arm, xlab="Arm", ylab="Change in Shannon Diversity")
```


## Evenness
```{r}
even_melted_metadata_filt <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("pielou_evenness"))
even_melted_metadata_filt_filt <- even_melted_metadata_filt %>% group_by(Patint_ID) %>% filter(n()==2) %>% mutate(slope=value[Week==12] - value[Week==0])

#remove all week12
even_melted_metadat_pairwise <- even_melted_metadata_filt_filt[-which(even_melted_metadata_filt_filt$Week==12),]

#faiths
boxplot(even_melted_metadat_pairwise$slope ~ even_melted_metadat_pairwise$arm, xlab="Arm", ylab="Change in Evenness")
```

## Richness
```{r}
rich_melted_metadata_filt <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("observed_features"))
rich_melted_metadata_filt_filt <- rich_melted_metadata_filt %>% group_by(Patint_ID) %>% filter(n()==2) %>% mutate(slope=value[Week==12] - value[Week==0])

#remove all week12
rich_melted_metadat_pairwise <- rich_melted_metadata_filt_filt[-which(rich_melted_metadata_filt_filt$Week==12),]

#faiths
boxplot(rich_melted_metadat_pairwise$slope ~ rich_melted_metadat_pairwise$arm, xlab="Arm", ylab="Change in Richness")
```
