---
title: "Beta Div. Analysis"
author: "Jacob T. Nearing"
date: "2/15/2021"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "light"
    downcute_theme: "chaos"
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(vegan)
library(ggplot2)
library(cowplot)
library(dplyr)
library(broom)
library(kableExtra)
library(reshape2)
library(ggpubr)
library(ggbeeswarm)

colors <- c('ferrous bisglycinate'="#a09342", 'ferrous sulfate'="#9673c5", 'placebo'="#ba6d75")

colors_weeks <- c('ferrous bisglycinate 0'="#a09342",
                  'ferrous bisglycinate 12'="#d5be2e",
                  'ferrous sulfate 0'="#9673c5",
                  'ferrous sulfate 12'="#6f0bf7",
                  'placebo 0'="#ba6d75",
                  'placebo 12'="#fa0922")
```

# Introduction

This analysis will examine whether there are overall changes in the microbial communities of individuals due to the use of different iron supplements.

```{r}
Metadata <- read.table("~/projects/IRON_SUP_EMMA/Data/Metadata/Iron_Metadata_21_03_09.txt", header=T, sep="\t", row.names=1)



bray_curt <- read.table("~/projects/IRON_SUP_EMMA/Data/diversity_5000/beta_raw/bray_curt/distance-matrix.tsv", header=T, row.names = 1)


weighted_uni <- read.table("~/projects/IRON_SUP_EMMA/Data/diversity_5000/beta_raw/weighted_unifrac/distance-matrix.tsv",
                           header=T, row.names = 1)
unweighted_uni <- read.table("~/projects/IRON_SUP_EMMA/Data/diversity_5000/beta_raw/unweighted_unifrac/distance-matrix.tsv",
                             header=T, row.names = 1)

Metadata_filt <- Metadata[rownames(weighted_uni),]

Metadata_filt <- Metadata_filt %>% group_by(Patint_ID) %>% filter(n()==2)

Metadata_week0 <- Metadata_filt %>% filter(Week==0) %>% 
  mutate(test=Patint_ID) %>%
  mutate(pair_dist_bray=bray_curt[paste("e",Patint_ID,sep=""), paste("b",Patint_ID,sep="")]) %>%
  mutate(pair_dist_wuni=weighted_uni[paste("e",Patint_ID,sep=""), paste("b",Patint_ID,sep="")]) %>%
  mutate(pair_dist_uuni=unweighted_uni[paste("e",Patint_ID,sep=""), paste("b",Patint_ID,sep="")]) %>% 
  ungroup()

sample_size <- Metadata_week0 %>% group_by(arm) %>% summarize(n())
sample_size

Metadata_week0 <- Metadata_week0 %>% left_join(data.frame(colors, arm=names(colors))) %>%
  left_join(sample_size) %>% mutate(axis_name=paste(arm,"\nn=",`n()`,sep=""))

```

# Examination of Pairwise distances {.tabset}

## Bray Curtis
```{r warning=F}
bray_distances <- ggplot(Metadata_week0, aes(x=axis_name, y=pair_dist_bray)) + 
  geom_boxplot(fill="white", width=0.1, outlier.shape = NA) + 
  geom_violin(alpha=0.3,aes(fill=colors)) +
  geom_quasirandom() +
  scale_fill_identity() + 
  xlab("Treatment Arm") +
  ylab("Bray-Curtis pairwise distance") + 
  theme_cowplot()

bray_distances
```

Add stats

```{r}
pair_bray_res <- compare_means(pair_dist_bray ~ axis_name, data=Metadata_week0, p.adjust.method = "holm", 
                               method="wilcox.test", ref.group = "placebo\nn=47") %>%
  mutate(y.position=c(1,1.1))

bray_distances + stat_pvalue_manual(pair_bray_res)
```


## Weighted UniFrac
```{r warning=F}
wuni_distances <- ggplot(Metadata_week0, aes(x=axis_name, y=pair_dist_wuni)) + 
  geom_boxplot(fill="white", width=0.1, outlier.shape = NA) + 
  geom_violin(alpha=0.3,aes(fill=colors)) +
  geom_quasirandom() +
  scale_fill_identity() + 
  xlab("Treatment Arm") +
  ylab("Weighted UniFrac pairwise distance") + 
  theme_cowplot()

wuni_distances
```

Add stats

```{r}
pair_wuni_res <- compare_means(pair_dist_wuni ~ axis_name, data=Metadata_week0, p.adjust.method = "holm",
                               method="wilcox.test", ref.group = "placebo\nn=47") %>%
  mutate(y.position=c(1,1.1))

wuni_distances + stat_pvalue_manual(pair_wuni_res)
```

## Unweighted UniFrac
```{r warning=F}
uuni_distances <-  ggplot(Metadata_week0, aes(x=axis_name, y=pair_dist_uuni)) + 
  geom_boxplot(fill="white", width=0.1, outlier.shape = NA) + 
  geom_violin(alpha=0.3,aes(fill=colors)) +
  geom_quasirandom() +
  scale_fill_identity() + 
  xlab("Treatment Arm") +
  ylab("Unweighted UniFrac pairwise distance") + 
  theme_cowplot()

uuni_distances
```

Add stats

```{r}
pair_uuni_res <- compare_means(pair_dist_uuni ~ axis_name, data=Metadata_week0, p.adjust.method="holm",
                              method="wilcox.test", ref.group="placebo\nn=47") %>%
  mutate(y.position=c(1,1.1))

pair_uuni_res

uuni_distances + stat_pvalue_manual(pair_uuni_res)
```

## Faceted pub plot

```{r}
plot_data <- Metadata_week0 %>% 
  select(axis_name, pair_dist_bray, pair_dist_wuni, pair_dist_uuni, colors) %>%
  melt()

plot_data$variable <- recode_factor(plot_data$variable, 
                                         pair_dist_wuni="Weighted UniFrac",
                                         pair_dist_bray="Bray Curtis",
                                         pair_dist_uuni="Unweighted UniFrac")

pair_dist <- plot_data %>%
  ggplot(aes(x=axis_name, y=value)) +
  geom_violin(alpha=0.3, aes(fill=colors)) +
  geom_boxplot(width=0.1, fill="white", outlier.shape = NA) +
  geom_quasirandom() +
  facet_wrap(~ variable) +
  scale_fill_identity() +
  xlab("") +
  ylab("Pairwise Distance") +
  theme_cowplot()

pair_dist
```


```{r}
pair_bray_res$variable <- "Bray Curtis"
pair_uuni_res$variable <- "Unweighted UniFrac"
pair_wuni_res$variable <- "Weighted UniFrac"

faceted_res <- do.call(rbind, list(pair_bray_res, pair_uuni_res, pair_wuni_res))


pair_dist + stat_pvalue_manual(faceted_res)
```


```{r}
Figure1_top_no_stats <- readRDS("~/GitHub_Repos/Cambodia_Iron_Supplement/Display_items/Intermediate_RDS_files/Figure1_top_no_stats.RDS")
Figure1_top_yes_stats <- readRDS("~/GitHub_Repos/Cambodia_Iron_Supplement/Display_items/Intermediate_RDS_files/Figure1_top_yes_stats.RDS")

Figure1_total_no_stats <- plot_grid(Figure1_top_no_stats, pair_dist, labels = c("A", "B"),
                                    nrow=2)
Figure1_total_no_stats

ggsave(filename = "~/GitHub_Repos/Cambodia_Iron_Supplement/Display_items/Main_Figures/Figure1.pdf",
       plot=Figure1_total_no_stats,
       width=10,
       height=9,
       dpi=600)

Figure1_total_yes_stats <- plot_grid(Figure1_top_yes_stats,
                                     pair_dist + stat_pvalue_manual(faceted_res), 
                                     labels=c("A", "B"),
                                     nrow = 2)
Figure1_total_yes_stats

ggsave(filename="~/GitHub_Repos/Cambodia_Iron_Supplement/Display_items/Main_Figures/Figure1_stats.pdf",
       plot=Figure1_total_yes_stats,
       width=10,
       height=9,
       dpi=600)

```

# Exploration of treatment data using Pcoa/PERMANOVA {.tabset}

First we will look at wether there is a significant interaction between week and arm to see if there is an effect on treatment. This is a second way of looking at a similar analysis as above and I would expect to receive similar results. The exepction to this is that the above analysis does not indicate the direction of the shift rather just the magnitude so this is another way to analysis the data taking into the account of the centroid and dispersion of the data.


## Bray Curtis {.tabset}

### All Weeks
```{r warning=F}
Metadata_filt <- Metadata_filt %>% 
  mutate(arm_week=paste(arm,Week, sep=" ")) %>%
  mutate(sample_name=paste(ifelse(Week==0, "b", "e"),Patint_ID,sep=""))

Metadata_filt$arm_week

bray_curt_filt <- bray_curt[Metadata_filt$sample_name, Metadata_filt$sample_name]

bray_curt_pcoa <- cmdscale(bray_curt_filt, eig=T, k=2)
comps_bray <- bray_curt_pcoa$eig/sum(bray_curt_pcoa$eig)*100

bray_plot_data <- data.frame(PC1=bray_curt_pcoa$points[,1], PC2=bray_curt_pcoa$points[,2], class=Metadata_filt$arm_week, sample=Metadata_filt$Patint_ID)

bray_plot_data <- left_join(bray_plot_data, data.frame(colors_weeks, class=names(colors_weeks)))
bray_plot_data$Week <- grepl("12", bray_plot_data$class)
bray_plot_data$Week <- ifelse(bray_plot_data$Week, "12", "0")

bray_plot <- ggplot(bray_plot_data, aes(x=PC1, y=PC2, color=colors_weeks, group=sample, shape=Week)) + 
  geom_point(size=3, alpha=0.8) + 
  theme_classic() + 
  #coord_fixed() + 
  xlab("PC1 5.7%") +
  ylab("PC2 4.7%") + 
  scale_color_identity("Sample Type", labels=gsub("_", " ",names(colors_weeks)), guide="legend", breaks=colors_weeks) +
  theme(plot.margin = margin(6, 0, 6, 0)) +
  ggtitle("Bray-Curtis")
  

bray_plot





#stop testing if the samples do not match up

stopifnot(identical(rownames(bray_curt_filt), rownames(Metadata_filt$sample_name)))

test <- adonis2(bray_curt_filt ~ arm*Week + Health_center, data=Metadata_filt, by="margin")
test
```

Calculate change in PC1 between time points for each arm.

```{r}

Bray_diff_data <- bray_plot_data %>% 
  group_by(sample) %>% 
  filter(n() == 2) %>% 
  mutate(PC1_diff=PC1[Week==12] - PC1[Week==0]) %>%
  filter(Week==12)

Bray_diff_data$class <- gsub(" 12", "",  Bray_diff_data$class)
Bray_diff_data$class

sample_size <- Bray_diff_data %>% group_by(class) %>% summarize(n())
sample_size

Bray_diff_data <- Bray_diff_data %>%
  left_join(sample_size) %>%
  left_join(data.frame(colors, class=names(colors))) %>%
  mutate(axis_name= paste0(class, "\n", "n=",`n()`))

Bray_diff_plot <- Bray_diff_data %>%
  ggplot(aes(x=axis_name, y=PC1_diff)) +
  geom_violin(alpha=0.3, aes(fill=colors)) +
  geom_boxplot(width=0.1, fill="white", outlier.shape = NA) +
  geom_quasirandom() +
  xlab("") +
  ylab("Change in PC1") + 
  ggtitle("Bray-Curtis") +
  scale_fill_identity() +
  theme_cowplot()
  
Bray_diff_plot
```


```{r}
pair_bray_pc1_res <- compare_means(PC1_diff ~ axis_name, data=Bray_diff_data, p.adjust.method="holm",
                              method="wilcox.test", ref.group="placebo\nn=47") %>%
  mutate(y.position=c(1,1.1))

pair_bray_pc1_res


Bray_diff_plot + stat_pvalue_manual(pair_bray_pc1_res)
```

Examine differences by week

```{r}

bray_plot_week <- ggplot(bray_plot_data, aes(x=PC1, y=PC2, color=Week, group=Week)) + 
  geom_point(size=3, alpha=0.8) + 
  theme_classic() + 
  #coord_fixed() + 
  xlab("PC1 5.7%") +
  ylab("PC2 4.7%") + 
  theme(plot.margin = margin(6, 0, 6, 0)) +
  ggtitle("Bray-Curtis") +
  stat_ellipse() +
  scale_color_manual(values=c("#a670b0", "#a58d48")) +
  theme_cowplot()
  

bray_plot_week


test <- adonis2(bray_curt_filt ~ Week + Health_center, data=Metadata_filt, by="margin")
test
```

## Weighted UniFrac {.tabset}

### All Weeks
```{r warning=F}

weighted_uni_filt <- weighted_uni[Metadata_filt$sample_name, Metadata_filt$sample_name]

weighted_uni_pcoa <- cmdscale(weighted_uni_filt, eig=T, k=2)

comps_wu <- weighted_uni_pcoa$eig/sum(weighted_uni_pcoa$eig)*100

wu_plot_data <- data.frame(PC1=weighted_uni_pcoa$points[,1], PC2=weighted_uni_pcoa$points[,2], class=Metadata_filt$arm_week, sample=Metadata_filt$Patint_ID)

wu_plot_data <- left_join(wu_plot_data, data.frame(colors_weeks, class=names(colors_weeks)))
wu_plot_data$Week <- grepl("12", wu_plot_data$class)
wu_plot_data$Week <- ifelse(wu_plot_data$Week, "12", "0")



wu_plot <- ggplot(wu_plot_data, aes(x=PC1, y=PC2, color=colors_weeks, group=sample, shape=Week)) + 
  geom_point(size=3, alpha=0.8) + 
  theme_classic() + 
  #coord_fixed() + 
  xlab("PC1 56.1%") +
  ylab("PC2 10.1%") + 
  scale_color_identity("Sample Type", labels=gsub("_", " ",names(colors_weeks)), guide="legend", breaks=colors_weeks) +
  theme(plot.margin = margin(6, 0, 6, 0)) +
  ggtitle("Weighted UniFrac")

wu_plot


stopifnot(identical(rownames(weighted_uni_filt), Metadata_filt$sample_name))

Adonis_test <- adonis2(weighted_uni_filt ~ arm*Week  + Health_center, data=Metadata_filt, by="margin")
Adonis_test
```


```{r}
wu_diff_data <- wu_plot_data %>% 
  group_by(sample) %>% 
  filter(n() == 2) %>% 
  mutate(PC1_diff=PC1[Week==12] - PC1[Week==0]) %>%
  filter(Week==12)

wu_diff_data$class <- gsub(" 12", "",  wu_diff_data$class)
wu_diff_data

wu_diff_data <- wu_diff_data %>%
  left_join(sample_size) %>%
  left_join(data.frame(colors, class=names(colors))) %>%
  mutate(axis_name= paste0(class, "\n", "n=",`n()`))
  
  
wu_diff_plot <- wu_diff_data %>%
  ggplot(aes(x=axis_name, y=PC1_diff)) +
  geom_violin(alpha=0.3, aes(fill=colors)) +
  geom_boxplot(width=0.1, fill="white", outlier.shape = NA) +
  geom_quasirandom() +
  xlab("") +
  ylab("Change in PC1") + 
  ggtitle("Weighted UniFrac") +
  scale_fill_identity() +
  theme_cowplot()
  


wu_diff_plot
```

```{r}
pair_wuni_pc1_res <- compare_means(PC1_diff ~ axis_name, data=wu_diff_data, p.adjust.method="holm",
                              method="wilcox.test", ref.group="placebo\nn=47") %>%
  mutate(y.position=c(1,1.1))

pair_wuni_pc1_res

wu_diff_plot + stat_pvalue_manual(pair_wuni_pc1_res)
```


```{r}
wu_plot_week <- ggplot(wu_plot_data, aes(x=PC1, y=PC2, color=Week, group=Week)) + 
  geom_point(size=3, alpha=0.8) + 
  theme_classic() + 
  #coord_fixed() + 
  xlab("PC1 56.1%") +
  ylab("PC2 10.1%") + 
  theme(plot.margin = margin(6, 0, 6, 0)) +
  ggtitle("Weighted UniFrac") +
  stat_ellipse() +
  scale_color_manual(values=c("#a670b0", "#a58d48")) +
  theme_cowplot()

wu_plot_week

test <- adonis2(weighted_uni_filt ~ Week + Health_center, data=Metadata_filt, by="margin")
test
```

## Unweighted UniFrac {.tabset}

### All Weeks
```{r warning=F}
unweighted_uni_filt <- unweighted_uni[Metadata_filt$sample_name, Metadata_filt$sample_name]

unweighted_uni_pcoa <- cmdscale(unweighted_uni_filt, eig=T, k=2)
comps_unwu <- unweighted_uni_pcoa$eig/sum(unweighted_uni_pcoa$eig)*100

unwu_plot_data <- data.frame(PC1=unweighted_uni_pcoa$points[,1], PC2=unweighted_uni_pcoa$points[,2], class=Metadata_filt$arm_week, sample=Metadata_filt$Patint_ID)


unwu_plot_data <- left_join(unwu_plot_data, data.frame(colors_weeks, class=names(colors_weeks)))
unwu_plot_data$Week <- grepl("12", unwu_plot_data$class)
unwu_plot_data$Week <- ifelse(unwu_plot_data$Week, "12", "0")


unwu_plot <- ggplot(unwu_plot_data, aes(x=PC1, y=PC2, color=colors_weeks, group=sample, shape=Week)) + 
  geom_point(size=3, alpha=0.8) + 
  theme_classic() + 
  #coord_fixed() + 
  xlab("PC1 9.5%") +
  ylab("PC2 5.6%") + 
  scale_color_identity("Sample Type", labels=gsub("_", " ",names(colors_weeks)), guide="legend", breaks=colors_weeks) +
  theme(plot.margin = margin(6, 0, 6, 0)) +
  ggtitle("Unweighted UniFrac")

unwu_plot


stopifnot(identical(rownames(unweighted_uni_filt), Metadata_filt$sample_name))

unweighted_test <- adonis2(unweighted_uni_filt ~ arm*Week + Health_center, by="margin", data=Metadata_filt)
unweighted_test

```

```{r}
unwu_diff_data <- unwu_plot_data %>% 
  group_by(sample) %>% 
  filter(n() == 2) %>% 
  mutate(PC1_diff=PC1[Week==12] - PC1[Week==0]) %>%
  filter(Week==12)

unwu_diff_data$class <- gsub(" 12", "",  unwu_diff_data$class)
unwu_diff_data

unwu_diff_data <- unwu_diff_data %>%
  left_join(sample_size) %>%
  left_join(data.frame(colors, class=names(colors))) %>%
  mutate(axis_name= paste0(class, "\n", "n=",`n()`))

unwu_diff_plot <- unwu_diff_data %>%
  ggplot(aes(x=axis_name, y=PC1_diff)) +
  geom_violin(alpha=0.3, aes(fill=colors)) +
  geom_boxplot(width=0.1, fill="white", outlier.shape = NA) +
  geom_quasirandom() +
  xlab("") +
  ylab("Change in PC1") + 
  ggtitle("Unweighted UniFrac") +
  scale_fill_identity() +
  theme_cowplot()
  


unwu_diff_plot
```

```{r}
pair_unwu_pc1_res <- compare_means(PC1_diff ~ axis_name, data=unwu_diff_data, p.adjust.method="holm",
                              method="wilcox.test", ref.group="placebo\nn=47") %>%
  mutate(y.position=c(1,1.1))

pair_unwu_pc1_res

unwu_diff_plot + stat_pvalue_manual(pair_unwu_pc1_res)
```

Examine by week

```{r}
unwu_plot_week <- ggplot(unwu_plot_data, aes(x=PC1, y=PC2, color=Week, group=Week)) + 
  geom_point(size=3, alpha=0.8) + 
  theme_classic() + 
  #coord_fixed() + 
  xlab("PC1 9.5%") +
  ylab("PC2 5.6%") + 
  theme(plot.margin = margin(6, 0, 6, 0)) +
  ggtitle("Unweighted UniFrac") +
  stat_ellipse() +
  scale_color_manual(values=c("#a670b0", "#a58d48")) +
  theme_cowplot()

unwu_plot_week

test <- adonis2(unweighted_uni_filt ~ Week + Health_center, data=Metadata_filt, by="margin")
test
```


## Publication Figure

```{r}
Figure2_top_temp <- plot_grid(bray_plot + theme(legend.position = "none"),
                         wu_plot + theme(legend.position = "none"),
                         unwu_plot + theme(legend.position = "none"),
                         nrow=1)
legend <- get_legend(bray_plot + theme(legend.box.margin = margin(0,0,0,12)))
legend

Figure2_top <- plot_grid(Figure2_top_temp, legend, rel_widths = c(3, .4))
Figure2_top


Figure2_bottom <- plot_grid(Bray_diff_plot, wu_diff_plot, unwu_diff_plot, nrow=1)
Figure2_bottom

Figure2 <- plot_grid(Figure2_top, Figure2_bottom, nrow=2)
Figure2

ggsave(filename = "~/GitHub_Repos/Cambodia_Iron_Supplement/Display_items/Main_Figures/Figure2.pdf",
       plot=Figure2,
       width=10,
       height=9,
       dpi=600)
```


```{r}
Supp_fig <- plot_grid(bray_plot_week, wu_plot_week, unwu_plot_week, nrow=1,
                      labels=c("A", "B", "C"))
Supp_fig
```
