---
title: "Differential Anundance Analysis"
author: "Jacob T. Nearing"
date: "3/2/2021"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "light"
    downcute_theme: "chaos"
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ALDEx2)
library(lme4)
library(lmerTest)
library(broom)
library(dplyr)
library(kableExtra)
library(corncob)
library(cowplot)
library(scales)
library(ggplot2)
library(ggbeeswarm)
```

```{r}
ANCOM_mod = function(feature_table, meta_data, struc_zero = NULL, main_var, p_adj_method = "BH", 
                 alpha = 0.05, adj_formula = NULL, rand_formula = NULL, fig=TRUE, ...){
  # OTU table transformation: 
  # (1) Discard taxa with structural zeros (if any); (2) Add pseudocount (1) and take logarithm.
  if (!is.null(struc_zero)) {
    num_struc_zero = apply(struc_zero, 1, sum)
    comp_table = feature_table[num_struc_zero == 0, ]
  }else{
    comp_table = feature_table
  }
  comp_table = log(as.matrix(comp_table) + 1)
  n_taxa = dim(comp_table)[1]
  taxa_id = rownames(comp_table)
  n_samp = dim(comp_table)[2]
  
  # Determine the type of statistical test and its formula.
  if (is.null(rand_formula) & is.null(adj_formula)) {
    # Basic model
    # Whether the main variable of interest has two levels or more?
    if (length(unique(meta_data%>%pull(main_var))) == 2) {
      # Two levels: Wilcoxon rank-sum test
      tfun = exactRankTests::wilcox.exact
    } else{
      # More than two levels: Kruskal-Wallis test
      tfun = stats::kruskal.test
    }
    # Formula
    tformula = formula(paste("x ~", main_var, sep = " "))
  }else if (is.null(rand_formula) & !is.null(adj_formula)) {
    # Model: ANOVA
    tfun = stats::aov
    # Formula
    tformula = formula(paste("x ~", main_var, "+", adj_formula, sep = " "))
  }else if (!is.null(rand_formula)) {
    # Model: Mixed-effects model
    tfun = nlme::lme
    # Formula
    if (is.null(adj_formula)) {
      # Random intercept model
      tformula = formula(paste("x ~", main_var))
    }else {
      # Random coefficients/slope model
      tformula = formula(paste("x ~", main_var, "+", adj_formula))
    }
  }
  
  
  if(grepl("\\*",main_var)){
    main_var <- gsub("\\*",":", main_var)
  }
  # Calculate the p-value for each pairwise comparison of taxa.
  p_data = matrix(NA, nrow = n_taxa, ncol = n_taxa)
  colnames(p_data) = taxa_id
  rownames(p_data) = taxa_id
  for (i in 1:(n_taxa - 1)) {
    # Loop through each taxon.
    # For each taxon i, additive log ratio (alr) transform the OTU table using taxon i as the reference.
    # e.g. the first alr matrix will be the log abundance data (comp_table) recursively subtracted 
    # by the log abundance of 1st taxon (1st column) column-wisely, and remove the first i columns since:
    # the first (i - 1) columns were calculated by previous iterations, and
    # the i^th column contains all zeros.
    alr_data = apply(comp_table, 1, function(x) x - comp_table[i, ]) 
    # apply(...) allows crossing the data in a number of ways and avoid explicit use of loop constructs.
    # Here, we basically want to iteratively subtract each column of the comp_table by its i^th column.
    alr_data = alr_data[, - (1:i), drop = FALSE]
    n_lr = dim(alr_data)[2] # number of log-ratios (lr)
    alr_data = cbind(alr_data, meta_data) # merge with the metadata
    
    # P-values
    if (is.null(rand_formula) & is.null(adj_formula)) {
      p_data[-(1:i), i] = apply(alr_data[, 1:n_lr, drop = FALSE], 2, function(x){
        tfun(tformula, data = data.frame(x, alr_data, check.names = FALSE))$p.value
        }
      ) 
    }else if (is.null(rand_formula) & !is.null(adj_formula)) {
      print(main_var)
      p_data[-(1:i), i] = apply(alr_data[, 1:n_lr, drop = FALSE], 2, function(x){
        fit = tfun(tformula, 
                   data = data.frame(x, alr_data, check.names = FALSE), 
                   na.action = na.omit)
        summary(fit)[[1]][main_var, "Pr(>F)"]
        }
      )
    }else if (!is.null(rand_formula)) {
      p_data[-(1:i), i] = apply(alr_data[, 1:n_lr, drop = FALSE], 2, function(x){
        fit = tfun(fixed = tformula, 
                   data = data.frame(x, alr_data, check.names = FALSE),
                   random = formula(rand_formula),
                   na.action = na.omit, ...)
        anova(fit)[main_var, "p-value"]
        }
      ) 
    }
  }
  # Complete the p-value matrix.
  # What we got from above iterations is a lower triangle matrix of p-values.
  p_data[upper.tri(p_data)] = t(p_data)[upper.tri(p_data)]
  diag(p_data) = 1 # let p-values on diagonal equal to 1
  # Multiple comparisons correction.
  q_data = apply(p_data, 2, function(x) p.adjust(x, method = p_adj_method))

  # Calculate the W statistic of ANCOM.
  # For each taxon, count the number of q-values < alpha.
  W = apply(q_data, 2, function(x) sum(x < alpha))
  
  # Organize outputs
  out_comp = data.frame(taxa_id, W, row.names = NULL, check.names = FALSE)
  # Declare a taxon to be differentially abundant based on the quantile of W statistic.
  # We perform (n_taxa - 1) hypothesis testings on each taxon, so the maximum number of rejections is (n_taxa - 1).
  out_comp = out_comp%>%mutate(detected_0.9 = ifelse(W > 0.9 * (n_taxa -1), TRUE, FALSE),
                               detected_0.8 = ifelse(W > 0.8 * (n_taxa -1), TRUE, FALSE),
                               detected_0.7 = ifelse(W > 0.7 * (n_taxa -1), TRUE, FALSE),
                               detected_0.6 = ifelse(W > 0.6 * (n_taxa -1), TRUE, FALSE))
  # Taxa with structural zeros are automatically declared to be differentially abundant
  if (!is.null(struc_zero)){
    out = data.frame(taxa_id = rownames(struc_zero), W = Inf, detected_0.9 = TRUE, 
                     detected_0.8 = TRUE, detected_0.7 = TRUE, detected_0.6 = TRUE, 
                     row.names = NULL, check.names = FALSE)
    out[match(taxa_id, out$taxa_id), ] = out_comp
  }else{
    out = out_comp
  }
  
  
  
  if(fig==FALSE){
    return(out)
  }
  # Draw volcano plot
  # Calculate clr
  clr_table = apply(feature_table, 2, clr)
  # Calculate clr mean difference
  eff_size = apply(clr_table, 1, function(y) 
    lm(y ~ x, data = data.frame(y = y, 
                                x = meta_data %>% pull(main_var),
                                check.names = FALSE))$coef[-1])
  
  if (is.matrix(eff_size)){
    # Data frame for the figure
    dat_fig = data.frame(taxa_id = out$taxa_id, t(eff_size), y = out$W, check.names = FALSE) %>% 
      mutate(zero_ind = factor(ifelse(is.infinite(y), "Yes", "No"), levels = c("Yes", "No"))) %>%
      gather(key = group, value = x, rownames(eff_size))
    # Replcace "x" to the name of covariate
    dat_fig$group = sapply(dat_fig$group, function(x) gsub("x", paste0(main_var, " = "), x))
    # Replace Inf by (n_taxa - 1) for structural zeros
    dat_fig$y = replace(dat_fig$y, is.infinite(dat_fig$y), n_taxa - 1)
    
    fig = ggplot(data = dat_fig) + aes(x = x, y = y) + 
      geom_point(aes(color = zero_ind)) + 
      facet_wrap(~ group) +
      labs(x = "CLR mean difference", y = "W statistic") +
      scale_color_discrete(name = "Structural zero", drop = FALSE) + 
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5), legend.position = "top",
            strip.background = element_rect(fill = "white"))
    fig  
  } else{
    # Data frame for the figure
    dat_fig = data.frame(taxa_id = out$taxa_id, x = eff_size, y = out$W) %>% 
      mutate(zero_ind = factor(ifelse(is.infinite(y), "Yes", "No"), levels = c("Yes", "No")))
    # Replace Inf by (n_taxa - 1) for structural zeros
    dat_fig$y = replace(dat_fig$y, is.infinite(dat_fig$y), n_taxa - 1)
    
    fig = ggplot(data = dat_fig) + aes(x = x, y = y) + 
      geom_point(aes(color = zero_ind)) + 
      labs(x = "CLR mean difference", y = "W statistic") +
      scale_color_discrete(name = "Structural zero", drop = FALSE) + 
      theme_bw() +
      theme(plot.title = element_text(hjust = 0.5), legend.position = "top")
    fig
  }
  
  res = list(out = out, fig = fig)
  return(res)
}
```

# Introductions

In this markdown we will explore whether there are specific taxa associated with the different iron supplement treatments used in this trail.

```{r}
remove_rare_features <- function( table , cutoff_pro=0.1, parallel=1 ) {
  if(cutoff_pro<=0){
    message("No filtering will be done due to cutoff_pro set to 0 or below")
    return(table)
  }
  row2keep <- c()
  cutoff <- ceiling( cutoff_pro * ncol(table) )
  if(parallel <= 1){
    for ( i in 1:nrow(table) ) {
      row_nonzero <- length( which( table[ i , ]  > 0 ) )
      if ( row_nonzero > cutoff ) {
        row2keep <- c( row2keep , i)
      }
    }
    return( table [ row2keep , , drop=F ])
  }else{
    `%dopar%` <- foreach::`%dopar%`
    cl <- parallel::makeCluster(parallel)
    doParallel::registerDoParallel(cl)
    message("Running with: ", foreach::getDoParWorkers(), " cores")

    res <- foreach::foreach(i = 1:nrow(table), .combine = c) %dopar% {
      row_nonzero <- length( which ( table[ i , ] > 0))
      if(row_nonzero > cutoff){
        i
      }else
        NULL
    }
    parallel::stopCluster(cl)
    return( table[res, , drop=F])

  }

}

```


```{r}
Metadata <- read.table("~/projects/IRON_SUP_EMMA/Data/Metadata/Iron_Metadata_21_03_09.txt", header=T, sep="\t", row.names=1)


### Load taxa table
Genus_table <- read.table("~/projects/IRON_SUP_EMMA/Data/TAXA_TABLES/genus_table/genus_table.tsv", sep="\t", header=T, row.names=1, check.names = F,
                          comment.char="", skip=1)

ASV_table <- read.table("~/projects/IRON_SUP_EMMA/Data/deblur_output/Raw_data/feature-table.tsv", sep="\t", header=T, row.names = 1, check.names = F,
                        comment.char = "", skip=1)




##only keep metadata from samples that sequenced sucessfully
Metadata_filt <- Metadata[colnames(Genus_table),]
Metadata_filt$arm_week <- paste(Metadata_filt$arm, Metadata_filt$Week, sep="_")
Metadata_filt$arm <- relevel(Metadata_filt$arm, ref="placebo")
Metadata_filt$Week <- as.factor(Metadata_filt$Week)

Metadata_filt <- Metadata_filt %>% group_by(Patint_ID) %>% filter(n()==2)
dim(Metadata_filt)

Metadata_filt <- Metadata_filt %>%  mutate(sample_name=paste(ifelse(Week==0, "b", "e"),Patint_ID,sep=""))

ASV_table <- ASV_table[,Metadata_filt$sample_name]
Genus_table <- Genus_table[,Metadata_filt$sample_name]

#use these post-hoc.
ASV_table_filt <- remove_rare_features(ASV_table, cutoff_pro = 0.05, parallel = 1)

Keep_ASV <- rownames(ASV_table_filt)

#use these post-hoc
Genus_table_filt <- remove_rare_features(Genus_table, cutoff_pro = 0.05, parallel = 1)

Keep_Genus <- rownames(Genus_table_filt)
```



```{r}
Genus_RA <- sweep(Genus_table, 2, colSums(Genus_table), '/')

flip_Genus_RA <- data.frame(t(Genus_RA))


ASV_RA <- sweep(ASV_table, 2, colSums(ASV_table), '/')

flip_ASV_RA <- data.frame(t(ASV_RA))


CLR_table_ASV <- data.frame(apply(ASV_table_filt + 1, 2, function(x){log(x) - mean(log(x))}))
flip_CLR_table_ASV <- data.frame(t(CLR_table_ASV))

### convert to CLR 
CLR_table <- data.frame(apply(Genus_table + 1, 2, function(x){log(x) - mean(log(x))}))
CLR_table_flip <- data.frame(t(CLR_table))


stopifnot(identical(rownames(CLR_table_flip), Metadata_filt$sample_name))
```


# ALDEx2 Analysis

## Aldex2 all glm interaction

This model is more appropriate and takes into the account the itneraction between being in different treatment groups and time since the start of treatment.

### Genus
```{r}
Metadata_filt$arm <- relevel(Metadata_filt$arm, ref="placebo")

MM <- model.matrix(~ arm + Week + arm*Week, data = Metadata_filt)
inter_genus_clr <- aldex.clr(Genus_table, MM, mc.samples = 128)
inter_genus_glm <- aldex.glm(inter_genus_clr, MM)


#filter to genera found in at least 5% of samples
inter_genus_glm <- inter_genus_glm[Keep_Genus,]


#readjust p-value correction
inter_genus_glm$`model.armferrous bisglycinate Pr(>|t|).BH` <- p.adjust(inter_genus_glm$`model.armferrous bisglycinate Pr(>|t|)`, method = "BH")
inter_genus_glm$`model.armferrous sulfate Pr(>|t|).BH` <- p.adjust(inter_genus_glm$`model.armferrous sulfate Pr(>|t|)`, method="BH")

min(inter_genus_glm$`model.armferrous bisglycinate Pr(>|t|).BH`)
min(inter_genus_glm$`model.armferrous sulfate Pr(>|t|).BH`)


saveRDS(inter_genus_glm, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/Genus_Aldex2.RDS")
```


Add in Health Center as a co-variate
```{r}
Metadata_filt$arm <- relevel(Metadata_filt$arm, ref="placebo")

MM <- model.matrix(~ arm + Week + Health_center + arm*Week, data = Metadata_filt)
inter_genus_clr <- aldex.clr(Genus_table, MM, mc.samples = 128)
inter_genus_glm <- aldex.glm(inter_genus_clr, MM)


#filter to genera found in at least 5% of samples
inter_genus_glm <- inter_genus_glm[Keep_Genus,]


#readjust p-value correction
inter_genus_glm$`model.armferrous bisglycinate Pr(>|t|).BH` <- p.adjust(inter_genus_glm$`model.armferrous bisglycinate Pr(>|t|)`, method = "BH")
inter_genus_glm$`model.armferrous sulfate Pr(>|t|).BH` <- p.adjust(inter_genus_glm$`model.armferrous sulfate Pr(>|t|)`, method="BH")

min(inter_genus_glm$`model.armferrous bisglycinate Pr(>|t|).BH`)
min(inter_genus_glm$`model.armferrous sulfate Pr(>|t|).BH`)


saveRDS(inter_genus_glm, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/Genus_Aldex2_HC.RDS")
```

No significant results

### ASV


With HC
```{r}
inter_ASV_clr <- aldex.clr(ASV_table, MM, mc.samples = 128)
inter_ASV_glm <- aldex.glm(inter_ASV_clr, MM)

inter_ASV_glm <- inter_ASV_glm[Keep_ASV,]

inter_ASV_glm$`model.armferrous bisglycinate Pr(>|t|).BH` <- p.adjust(inter_ASV_glm$`model.armferrous bisglycinate Pr(>|t|)`, "BH")
inter_ASV_glm$`model.armferrous sulfate Pr(>|t|).BH` <- p.adjust(inter_ASV_glm$`model.armferrous sulfate Pr(>|t|)`, "BH")


min(inter_ASV_glm$`model.armferrous bisglycinate Pr(>|t|).BH`)
min(inter_ASV_glm$`model.armferrous sulfate Pr(>|t|).BH`)



saveRDS(inter_ASV_glm, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/ASV_Aldex2_HC.RDS")

```


Without HC
```{r}
MM <- model.matrix(~ arm + Week + arm*Week, data = Metadata_filt)
inter_ASV_clr <- aldex.clr(ASV_table, MM, mc.samples = 128)
inter_ASV_glm <- aldex.glm(inter_ASV_clr, MM)

inter_ASV_glm <- inter_ASV_glm[Keep_ASV,]

inter_ASV_glm$`model.armferrous bisglycinate Pr(>|t|).BH` <- p.adjust(inter_ASV_glm$`model.armferrous bisglycinate Pr(>|t|)`, "BH")
inter_ASV_glm$`model.armferrous sulfate Pr(>|t|).BH` <- p.adjust(inter_ASV_glm$`model.armferrous sulfate Pr(>|t|)`, "BH")


min(inter_ASV_glm$`model.armferrous bisglycinate Pr(>|t|).BH`)
min(inter_ASV_glm$`model.armferrous sulfate Pr(>|t|).BH`)



saveRDS(inter_ASV_glm, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/ASV_Aldex2.RDS")

```

No Significant results.

# Corncob Examination

Corncob is a another tool used to identify microbes that might be associated with a specific group of interest.

## All individuals

### Gensus level analysis {.tabset}
```{r}
Metadata_filt$arm <- relevel(Metadata_filt$arm, ref="placebo")
Metadata_filt$Week <- factor(Metadata_filt$Week, levels = c(0,12))

Metadata_filt <- data.frame(Metadata_filt)
rownames(Metadata_filt) <- Metadata_filt$sample_name

otu <- phyloseq::otu_table(Genus_table, taxa_are_rows = T)
sampledata <- phyloseq::sample_data(Metadata_filt)
phylo <- phyloseq::merge_phyloseq(otu, sampledata)

apply(phylo@otu_table, 1, FUN = function(x) wilcox.test(x ~ phylo@sam_data$Week))

wilcox.test(phylo@otu_table ~ phylo@sam_data$Week)

results_interact <- differentialTest(formula = ~ arm + Week + arm:Week,
                                     phi.formula = ~ arm + Week + arm:Week,
                                     phi.formula_null = ~ arm + Week + arm:Week,
                                     formula_null = ~ arm + Week,
                                     test="Wald", 
                                     data=phylo,
                                     boot=F,
                                     fdr_cutoff = 0.1)

saveRDS(results_interact, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/Genus_corncob.RDS")

pvals <- results_interact$p
pvals_filt <- pvals[names(pvals) %in% Keep_Genus]
qvals_filt <- p.adjust(pvals_filt, "BH")

min(qvals_filt, na.rm = T)

which(qvals_filt < 0.1)


plot(results_interact)
```

```{r}
stopifnot(identical(rownames(Metadata_filt), rownames(flip_Genus_RA)))
```


```{r}
sloop_function <- function(sloop_val){
  
  if(sloop_val==0){
    return("Constant")
  }else if(sloop_val > 0){
    return("Decrease")
  }else
    return("Increase")
  
}

```

#### Enterococcus
```{r}
comb_data <- cbind(Metadata_filt, flip_Genus_RA)


melted_metadata_filt <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("d__Bacteria.p__Firmicutes.c__Bacilli.o__Lactobacillales.f__Enterococcaceae.g__Enterococcus"))

melted_metadata_filt_filt <- melted_metadata_filt %>% 
  group_by(Patint_ID) %>% 
  filter(n()==2) %>% 
  mutate(slope=sloop_function(value[Week==0] - value[Week==12]))


diff_plot_enterococcus <- ggplot(melted_metadata_filt_filt, aes(x=as.character(Week), y=log(value+0.00001, base=10))) + 
  geom_violin() + 
  geom_boxplot(aes(fill=Week), alpha=0.2, width=0.1) +
  geom_point(alpha=0.5) +
  geom_line(aes(color=slope, group=Patint_ID), alpha=.5) + facet_grid(~ arm) +
  ylab("log10(Enterococcus RA.)") + xlab("Week of treatment") + 
  scale_color_brewer(name="Direction of change", labels=c("Constant", "Decrease", "Increase"), palette="Dark2") + theme_cowplot() + 
  guides(fill="none")

diff_plot_enterococcus

table_data_entero <- table(melted_metadata_filt_filt$slope, melted_metadata_filt_filt$arm)/2
enterococcous_table <- gridExtra::tableGrob(table_data_entero)
enterococcous_table <- ggdraw(enterococcous_table)
enterococcous_table
```

#### Weissella
```{r}
comb_data <- cbind(Metadata_filt, flip_Genus_RA)

melted_metadata_weissella <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("d__Bacteria.p__Firmicutes.c__Bacilli.o__Lactobacillales.f__Leuconostocaceae.g__Weissella"))

melted_metadata_weissella_weissella <- melted_metadata_weissella %>% 
  group_by(Patint_ID) %>% 
  filter(n()==2) %>% 
  mutate(slope=sloop_function(value[Week==0] - value[Week==12]))


weissella_plot <- ggplot(melted_metadata_weissella_weissella, aes(x=as.character(Week), y=log(value + 0.00001, base = 10))) + 
  geom_violin() +
  geom_boxplot(aes(fill=Week), alpha=0.2, width=0.1) +
  geom_point(alpha=0.5) + 
  geom_line(aes(color=slope, group=Patint_ID), alpha=.5) + facet_grid(~ arm) +
  ylab("log10(Weissella RA.)") + xlab("Week of treatment") + scale_color_brewer(name="Direction of change", labels=c("Constant", "Decrease", "Increase"), palette = "Dark2") + 
  theme_cowplot() + guides(fill='none')


weissella_plot

weissella_table <- gridExtra::tableGrob(table(melted_metadata_weissella_weissella$slope, melted_metadata_weissella_weissella$arm)/2)
weissella_table <- ggdraw(weissella_table)
weissella_table
```

#### Escherichia-Shigella
```{r}


melted_metadata_escherichia <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("d__Bacteria.p__Proteobacteria.c__Gammaproteobacteria.o__Enterobacterales.f__Enterobacteriaceae.g__Escherichia.Shigella"))

melted_metadata_escherichia_escherichia <- melted_metadata_escherichia %>% 
  group_by(Patint_ID) %>% 
  filter(n()==2) %>% 
  mutate(slope=sloop_function(value[Week==0] - value[Week==12]))



escherichia_plot <- ggplot(melted_metadata_escherichia_escherichia, aes(x=as.character(Week), y=log(value + 0.000001, base=10))) + 
  geom_violin() +
  geom_boxplot(aes(fill=Week), alpha=0.2, width=0.1) +
  geom_point(alpha=0.4) + 
  geom_line(aes(color=slope, group=Patint_ID), alpha=.5) + facet_grid(~ arm) +
  ylab("log10(Escherichia-Shigella RA.)") + xlab("Week of treatment") +
  scale_color_brewer(name="Direction of change", labels=c("Decrease", "Increase"), palette = "Dark2") + 
  theme_cowplot() + guides(fill='none')

escherichia_plot

melted_metadata_escherichia_escherichia$slope <- factor(melted_metadata_escherichia_escherichia$slope, levels = c("Constant", "Decrease", "Increase"))


escherichia_table <- gridExtra::tableGrob(table(melted_metadata_escherichia_escherichia$slope, melted_metadata_escherichia_escherichia$arm)/2)
escherichia_table <- ggdraw(escherichia_table)
escherichia_table
```

### Publication figure

```{r}
Figure3 <- plot_grid(diff_plot_enterococcus + theme(legend.position = "none"), 
                     weissella_plot + theme(legend.position = "none"), 
                     escherichia_plot + theme(legend.position = "none"), nrow=1)
Figure3_bottom <- plot_grid(enterococcous_table, weissella_table, escherichia_table, nrow=1)
Figure3_bottom

Figure3_temp <- plot_grid(Figure3, Figure3_bottom, nrow=2,
                          rel_heights = c(1,.25))
Figure3_temp

legend <- get_legend(diff_plot_enterococcus)
legend
Figure3_final <- plot_grid(Figure3_temp, legend, rel_widths = c(3,.4))
Figure3_final

ggsave("~/GitHub_Repos/Cambodia_Iron_Supplement/Display_items/Main_Figures/Figure3.pdf", plot=Figure3_final,
       width=9, height=11, units = "in")
```


### Genus with HC
```{r}
results_interact_HC <- differentialTest(formula = ~ arm + Week + arm:Week + Health_center,
                                     phi.formula = ~ arm + Week + arm:Week + Health_center,
                                     phi.formula_null = ~ arm + Week + arm:Week + Health_center,
                                     formula_null = ~ arm + Week + Health_center,
                                     test="Wald", 
                                     data=phylo,
                                     boot=F,
                                     fdr_cutoff = 0.1)


pvals <- results_interact_HC$p
pvals_filt <- pvals[names(pvals) %in% Keep_Genus]
qvals_filt <- p.adjust(pvals_filt, "BH")

min(qvals_filt, na.rm = T)

which(qvals_filt < 0.22)

saveRDS(results_interact_HC, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/Genus_corncob_HC.RDS")
```

Doesn't pass significance after apply Health_Center as a control.




### ASV analysis {.tabset}
```{r}
otu <- phyloseq::otu_table(ASV_table, taxa_are_rows = T)
sampledata <- phyloseq::sample_data(Metadata_filt)
phylo <- phyloseq::merge_phyloseq(otu, sampledata)

results_ASV_interact <- differentialTest(formula = ~ arm + Week + arm:Week,
                                     phi.formula = ~ arm + Week + arm:Week,
                                     phi.formula_null = ~ arm + Week + arm:Week,
                                     formula_null = ~ arm + Week,
                                     test="Wald", data=phylo,
                                     boot=F,
                                     fdr_cutoff = 0.1)


saveRDS(results_ASV_interact, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/ASV_corncob.RDS")


pvals <- results_ASV_interact$p
pvals_filt <- pvals[names(pvals) %in% Keep_ASV]
qvals_filt <- p.adjust(pvals_filt, "BH")
which.min(qvals_filt)

```


### ASV analysis with HC {.tabset}

```{r}
results_ASV_interact_HC <- differentialTest(formula = ~ arm + Week + arm:Week + Health_center,
                                     phi.formula = ~ arm + Week + arm:Week + Health_center,
                                     phi.formula_null = ~ arm + Week + arm:Week + Health_center,
                                     formula_null = ~ arm + Week + Health_center,
                                     test="Wald", data=phylo,
                                     boot=F,
                                     fdr_cutoff = 0.1)

saveRDS(results_ASV_interact_HC, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/ASV_corncob_HC.RDS")



pvals <- results_ASV_interact_HC$p
pvals_filt <- pvals[names(pvals) %in% Keep_ASV]
qvals_filt <- p.adjust(pvals_filt, "BH")
which.min(qvals_filt)
```

### 83e3b434607e9ea107c21d0c9640cee2

```{r}
comb_data <- cbind(Metadata_filt, flip_ASV_RA)

melted_metadata_83e3 <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("X83e3b434607e9ea107c21d0c9640cee2"))



melted_metadata_83e3_83e3 <- melted_metadata_83e3 %>% 
  group_by(Patint_ID) %>% filter(n()==2) %>% 
  mutate(slope=sloop_function(value[Week==0] - value[Week==12]))



X83e3_plot <- ggplot(melted_metadata_83e3_83e3, aes(x=as.character(Week), y=log(value + 0.000001, base=10))) + 
  geom_violin() +
  geom_boxplot(aes(fill=Week), alpha=0.2, width=0.1) +
  geom_point(alpha=0.4) + 
  geom_line(aes(color=slope, group=Patint_ID), alpha=.5) + facet_grid(~ arm) +
  ylab("log10(ASV: C. perfringens RA.)") + xlab("Week of treatment") +
  scale_color_brewer(name="Direction of change", labels=c("Constant","Decrease", "Increase"), palette = "Dark2") + 
  theme_cowplot() + guides(fill='none')

X83e3_plot

X83e3_table <- gridExtra::tableGrob(table(melted_metadata_83e3_83e3$slope, melted_metadata_83e3_83e3$arm)/2)
X83e3_table <- ggdraw(X83e3_table)
X83e3_table

```

### b623615586e4850b4d1e433eabb8cec6

```{r}
melted_metadata_b623 <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("b623615586e4850b4d1e433eabb8cec6"))



melted_metadata_b623_b623 <- melted_metadata_b623 %>% 
  group_by(Patint_ID) %>% filter(n()==2) %>% mutate(slope=sloop_function(value[Week==0] - value[Week==12]))



b623_plot <- ggplot(melted_metadata_b623_b623, aes(x=as.character(Week), y=log(value + 0.000001, base=10))) + 
  geom_violin() +
  geom_boxplot(aes(fill=Week), alpha=0.2, width=0.1) +
  geom_point(alpha=0.4) + 
  geom_line(aes(color=slope, group=Patint_ID), alpha=.5) + facet_grid(~ arm) +
  ylab("log10(E. coprostanoligenes RA.)") + xlab("Week of treatment") +
  scale_color_brewer(name="Direction of change", labels=c("Constant","Decrease", "Increase"), palette = "Dark2") + 
  theme_cowplot() + guides(fill='none')

b623_plot


b623_table <- gridExtra::tableGrob(table(melted_metadata_b623_b623$slope, melted_metadata_b623_b623$arm)/2)
b623_table <- ggdraw(b623_table)
b623_table
```


### 102807be929d5bfa1e0cd78ebaf5125f

```{r}
melted_metadata_X1028 <- reshape2::melt(comb_data, id.vars=c("Patint_ID", "Week", "arm"), measure.vars=c("X102807be929d5bfa1e0cd78ebaf5125f"))



melted_metadata_X1028_X1028 <- melted_metadata_X1028 %>% 
  group_by(Patint_ID) %>% filter(n()==2) %>% mutate(slope=sloop_function(value[Week==0] - value[Week==12]))



X1028_plot <- ggplot(melted_metadata_X1028_X1028, aes(x=as.character(Week), y=log(value + 0.000001, base=10))) + 
  geom_violin() +
  geom_boxplot(aes(fill=Week), alpha=0.2, width=0.1) +
  geom_point(alpha=0.4) + 
  geom_line(aes(color=slope, group=Patint_ID), alpha=.5) + facet_grid(~ arm) +
  ylab("log10(ASV: Bacteroides RA.)") + xlab("Week of treatment") +
  scale_color_brewer(name="Direction of change", labels=c("Constant","Decrease", "Increase"), palette = "Dark2") + 
  theme_cowplot() + guides(fill='none')

X1028_plot


X1028_table <- gridExtra::tableGrob(table(melted_metadata_X1028_X1028$slope, melted_metadata_X1028_X1028$arm)/2)
X1028_table <- ggdraw(X1028_table)
X1028_table
```

Supplemental Fig

```{r}
Figure_sup <- plot_grid(X83e3_plot + theme(legend.position = "none"), 
                     X1028_plot + theme(legend.position = "none"), 
                     b623_plot + theme(legend.position = "none"), nrow=1)

Figure_sup_bottom <- plot_grid(X83e3_table, X1028_table, b623_table, nrow=1)
Figure_sup_bottom

Figure_sup_temp <- plot_grid(Figure_sup, Figure_sup_bottom, nrow=2,
                          rel_heights = c(1,.25))
Figure_sup_temp

legend <- get_legend(X83e3_plot)
legend


Figure_sup_final <- plot_grid(Figure_sup_temp, legend, rel_widths = c(3,.4))
Figure_sup_final

```


# Maaslin2

## ASV
```{r}
library(Maaslin2)

ASV_table_flip <- data.frame(t(ASV_table), check.rows = F, check.names = F)

identical(rownames(Metadata_filt), rownames(ASV_table_flip))

Metadata_filt$ferrous_bis_12 <- (Metadata_filt$arm=="ferrous bisglycinate") * (Metadata_filt$Week=="12")
Metadata_filt$ferrous_sul_12 <- (Metadata_filt$arm=="ferrous sulfate") * (Metadata_filt$Week=="12")


fit_data <- Maaslin2(input_data = ASV_table_flip,
                     input_metadata = Metadata_filt,
                     output="~/projects/IRON_SUP_EMMA/Data/Maaslin2_res_int",
                     transform= "AST",
                     fixed_effects = c("arm", "Week", "ferrous_bis_12", "ferrous_sul_12"))

ferrous_bis_12_fit <- fit_data$results[grep("ferrous_bis_12", fit_data$results$name),]
ferrous_sul_12_fit <- fit_data$results[grep("ferrous_sul_12", fit_data$results$name),]

ferrous_bis_12_fit <- ferrous_bis_12_fit[ferrous_bis_12_fit$feature %in% Keep_ASV,]
ferrous_bis_12_fit$qval_filt <- p.adjust(ferrous_bis_12_fit$pval, "BH")
min(ferrous_bis_12_fit$qval_filt)


ferrous_sul_12_fit <- ferrous_sul_12_fit[ferrous_sul_12_fit$feature %in% Keep_ASV,]
ferrous_sul_12_fit$qval_filt <- p.adjust(ferrous_sul_12_fit$pval, "BH")
ferrous_sul_12_fit$qval_filt

```
No Hits.

## ASV with HC

```{r}
fit_data <- Maaslin2(input_data = ASV_table_flip,
                     input_metadata = Metadata_filt,
                     output="~/projects/IRON_SUP_EMMA/Data/Maaslin2_res_int",
                     transform= "AST",
                     fixed_effects = c("arm", "Week", "Health_center", "ferrous_bis_12", "ferrous_sul_12"))

ferrous_bis_12_fit <- fit_data$results[grep("ferrous_bis_12", fit_data$results$name),]
ferrous_sul_12_fit <- fit_data$results[grep("ferrous_sul_12", fit_data$results$name),]

ferrous_bis_12_fit <- ferrous_bis_12_fit[ferrous_bis_12_fit$feature %in% Keep_ASV,]
ferrous_bis_12_fit$qval_filt <- p.adjust(ferrous_bis_12_fit$pval, "BH")
min(ferrous_bis_12_fit$qval_filt)


ferrous_sul_12_fit <- ferrous_sul_12_fit[ferrous_sul_12_fit$feature %in% Keep_ASV,]
ferrous_sul_12_fit$qval_filt <- p.adjust(ferrous_sul_12_fit$pval, "BH")
ferrous_sul_12_fit$qval_filt

```


## Genus

```{r}

Genus_table_flip <- data.frame(t(Genus_table), check.rows = F, check.names = F)

identical(rownames(Genus_table_flip), rownames(Metadata_filt))

fit_data_genus <- Maaslin2(input_data = Genus_table_flip,
                     input_metadata = Metadata_filt,
                     output="~/projects/IRON_SUP_EMMA/Data/Maaslin2_res_genus",
                     transform="AST",
                     fixed_effects=c("arm","Week", "ferrous_bis_12", "ferrous_sul_12"))


genus_ferrous_bis_12 <- fit_data_genus$results[grep("ferrous_bis_12", fit_data_genus$results$name),]
genus_ferrous_sul_12 <- fit_data_genus$results[grep("ferrous_sul_12", fit_data_genus$results$name),]


genus_ferrous_bis_12 <- genus_ferrous_bis_12[genus_ferrous_bis_12$feature %in% Keep_Genus,]
genus_ferrous_bis_12$qval_filt <- p.adjust(genus_ferrous_bis_12$pval, "BH")
genus_ferrous_bis_12$qval_filt


genus_ferrous_sul_12 <- genus_ferrous_sul_12[genus_ferrous_sul_12$feature %in% Keep_Genus,]
genus_ferrous_sul_12$qval_filt <- p.adjust(genus_ferrous_sul_12$pval, "BH")
genus_ferrous_sul_12$qval_filt

```
No Hits.


## Genus with HC

```{r}
fit_data_genus <- Maaslin2(input_data = Genus_table_flip,
                     input_metadata = Metadata_filt,
                     output="~/projects/IRON_SUP_EMMA/Data/Maaslin2_res_genus",
                     transform="AST",
                     fixed_effects=c("arm","Week", "Health_center", "ferrous_bis_12", "ferrous_sul_12"))


genus_ferrous_bis_12 <- fit_data_genus$results[grep("ferrous_bis_12", fit_data_genus$results$name),]
genus_ferrous_sul_12 <- fit_data_genus$results[grep("ferrous_sul_12", fit_data_genus$results$name),]


genus_ferrous_bis_12 <- genus_ferrous_bis_12[genus_ferrous_bis_12$feature %in% Keep_Genus,]
genus_ferrous_bis_12$qval_filt <- p.adjust(genus_ferrous_bis_12$pval, "BH")
genus_ferrous_bis_12$qval_filt


genus_ferrous_sul_12 <- genus_ferrous_sul_12[genus_ferrous_sul_12$feature %in% Keep_Genus,]
genus_ferrous_sul_12$qval_filt <- p.adjust(genus_ferrous_sul_12$pval, "BH")
genus_ferrous_sul_12$qval_filt

```

# ANCOM-II

## ASV
```{r}
source("~/GitHub_Repos/Clean_Hackathon/Pipeline_scripts/Ancom2_Script/ancom_v2.1.R")

Metadata_filt$samp_id <- rownames(Metadata_filt)

prepro_asv <- feature_table_pre_process(feature_table=ASV_table,
                                        meta_data = Metadata_filt, sample_var = 'samp_id', group_var = NULL, out_cut=0.05, zero_cut = 0.9, lib_cut = 1000, neg_lb=FALSE)

res <- ANCOM_mod(feature_table = prepro_asv$feature_table, meta_data = prepro_asv$meta_data, struc_zero = prepro_asv$structure_zeros, main_var = "arm*Week",
             p_adj_method = "BH", alpha=0.1, adj_formula = "arm + Week", rand_formula = NULL, fig=FALSE)

saveRDS(res, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/ASV_ancomII.RDS")

```

## ASV with HC

```{r}
res <- ANCOM_mod(feature_table = prepro_asv$feature_table, meta_data = prepro_asv$meta_data, struc_zero = prepro_asv$structure_zeros, main_var = "arm*Week",
             p_adj_method = "BH", alpha=0.1, adj_formula = "arm + Week + Health_center", rand_formula = NULL, fig=FALSE)


saveRDS(res, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/ASV_ancomII_HC.RDS")
```

## Genus

```{r}

prepro_genus <- feature_table_pre_process(feature_table=Genus_table,
                                        meta_data = Metadata_filt, sample_var = 'samp_id', group_var = NULL, out_cut=0.05, zero_cut = 0.9, lib_cut = 1000, neg_lb=FALSE)


res_genus <- ANCOM_mod(feature_table = prepro_genus$feature_table, meta_data = prepro_genus$meta_data, 
                       struc_zero = prepro_genus$structure_zeros, main_var = "arm*Week",
             p_adj_method = "BH", alpha=0.1, adj_formula = "arm + Week", rand_formula = NULL, fig=FALSE)


saveRDS(res_genus, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/Genus_ancomII.RDS")

```

## Genus with HC

```{r}

res_genus <- ANCOM_mod(feature_table = prepro_genus$feature_table, meta_data = prepro_genus$meta_data, 
                       struc_zero = prepro_genus$structure_zeros, main_var = "arm*Week",
             p_adj_method = "BH", alpha=0.1, adj_formula = "arm + Week + Health_center", rand_formula = NULL, fig=FALSE)

saveRDS(res_genus, "~/GitHub_Repos/Cambodia_Iron_Supplement/DA_results/Genus_ancomII_HC.RDS")
```