---
title: "QPCR_analysis"
author: "Jacob T. Nearing"
date: "09/06/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(lme4)
library(broom)
library(broom.mixed)
library(ggplot2)
library(ggpubr)
library(cowplot)
library(ggbeeswarm)


colors <- c('ferrous bisglycinate'="#a09342", 'ferrous sulfate'="#9673c5", 'placebo'="#ba6d75")
```

```{r cars}
#load in metadata
metadata <- read.table("~/projects/IRON_SUP_EMMA/Data/Metadata/Iron_Metadata_21_03_09.txt", header=T, sep="\t", row.names=1)

#load in qpcr data for baseline (week 0 samples)
qpcr_data_inital <- as.data.frame(readxl::read_excel("~/projects/IRON_SUP_EMMA/qPCR_data/bfpa and eae results_CIT2.xlsx", sheet=1))
qpcr_data_inital$individual <- qpcr_data_inital$ID
qpcr_data_inital$ID <- paste0("b",qpcr_data_inital$ID)

#load in qpcr data for endpoint (week 12 samples)
qpcr_data_end <- as.data.frame(readxl::read_excel("~/projects/IRON_SUP_EMMA/qPCR_data/bfpa and eae results_CIT2.xlsx", sheet=2))
qpcr_data_end$individual <- qpcr_data_end$ID
qpcr_data_end$ID <- paste0("e", qpcr_data_end$ID)

#bind them together
qpcr_data_full <- rbind(qpcr_data_end, qpcr_data_inital)
row.names(qpcr_data_full) <- qpcr_data_full$ID

#filter to samples that have both a baseline and endpoint sample reading.
metadata_filt <- metadata[rownames(qpcr_data_full),]
metadata_filt <- metadata_filt %>% group_by(Patint_ID) %>% filter(n()==2) %>% ungroup()

qpcr_data_full <- qpcr_data_full %>% filter(individual %in% metadata_filt$Patint_ID)


full_data <- cbind(metadata_filt, qpcr_data_full)

# set reference groups
full_data$`bfpA +/-` <- factor(full_data$`bfpA +/-`)
full_data$`eae +/-` <- factor(full_data$`eae +/-`)


full_data$Week <- factor(full_data$Week)

full_data$arm <- factor(full_data$arm, levels=c("placebo", "ferrous bisglycinate", "ferrous sulfate"))


## load in standards data
bfpa_standard <- as.data.frame(readxl::read_excel("~/projects/IRON_SUP_EMMA/qPCR_data/standard curves - bfpa eae.xlsx", sheet=1))
eae_standard <- as.data.frame(readxl::read_excel("~/projects/IRON_SUP_EMMA/qPCR_data/standard curves - bfpa eae.xlsx", sheet=2))
```


# Presence Abscene Analysis

## Summary tables

### bfpa
```{r}
grp_numbers <- full_data %>% group_by(arm,Week,arm) %>% summarize(cnt=n())
grp_numbers

bfpa <- full_data %>% group_by(arm, Week, `bfpA +/-`) %>% summarize(cnt=n())
bfpa


#get % postive in each grouping
bfpa$cnt[1:4] <- bfpa$cnt[1:4]/56 * 100
bfpa$cnt[5:8] <- bfpa$cnt[5:8]/53 * 100
bfpa$cnt[9:12] <- bfpa$cnt[9:12]/57*100

bfpa_pos <- bfpa %>% filter(`bfpA +/-`==1)
bfpa_pos
```

### eae
```{r}
eae <- full_data %>% group_by(arm, Week, `eae +/-`) %>% summarize(cnt=n())
eae

eae$cnt[1:4] <- eae$cnt[1:4]/56 * 100
eae$cnt[5:8] <- eae$cnt[5:8]/53 * 100
eae$cnt[9:12] <- eae$cnt[9:12]/57 * 100

eae_pos <- eae %>% filter(`eae +/-`==1)
eae_pos
```

## Mixed effects logistic regression

### eae presence abscene

```{r}
eae_melr <- glmer(`eae +/-` ~ Week*arm + (1|Patint_ID), data=full_data, family="binomial")
eae_melr %>% tidy(conf.int=T)
```

### bfpa presence abscene
```{r}
bfpa_melr <- glmer(`bfpA +/-` ~ Week*arm + (1|Patint_ID), data=full_data, family="binomial")
bfpa_melr %>% tidy(conf.int=T)
```


# Abundance analysis


## eae

### standard curve

```{r}
colnames(eae_standard)[7] <- "eae Cq"

eae_standard$`eae Cq` <- as.numeric(eae_standard$`eae Cq`)
eae_standard$`eae Cq`


eae_lm  <- lm(`Log Starting Quantity` ~ `eae Cq`, data=eae_standard)
summary(eae_lm)

plot(eae_standard$`Log Starting Quantity` ~ eae_standard$`eae Cq`, xlab="Cq Value", ylab="Log Starting Quantity (ng)") 
abline(eae_lm)
```

Use standard curve to calculate values in true dataset

We will set log quantities to the detection limit for those samples that did not return a cq value.
```{r}
full_data$`eae Cq` <- as.numeric(full_data$`eae Cq`)
full_data$`eae Cq`


full_data$log_eae_ng <- predict(eae_lm, full_data)
full_data$log_eae_ng

full_data$log_eae_ng[which(is.na(full_data$log_eae_ng))] <- -4.6197888

plot(full_data$log_eae_ng)
```

### Compare change in log_eae_ng between week 0 and week 12

```{r}
eae_change <- full_data  %>% group_by(Patint_ID) %>% arrange(Week) %>%
  mutate("Change in log eae ng"=log_eae_ng - lag(log_eae_ng, default=NA)) %>% filter(Week==12) %>% ungroup

sample_size <- eae_change %>% group_by(arm) %>% summarize(n())
sample_size

eae_change <- eae_change %>% 
  left_join(sample_size) %>%
  left_join(data.frame(colors, arm=names(colors))) %>%
  mutate(myaxis=paste0(arm, "\n", "n=",`n()`))

eae_change_plot <- eae_change %>% 
  ggplot(aes(x=myaxis, y=`Change in log eae ng`)) +
  geom_violin(alpha=0.3, aes(fill=colors)) +
  geom_boxplot(fill="white", width=0.1, outlier.shape = NA) +
  geom_quasirandom() +
  xlab("") +
  ylab("Change in log(eae qauntity, ng)") +
  scale_fill_identity() +
  theme_cowplot()
  
eae_change_plot  

ggsave("~/GitHub_Repos/Cambodia_Iron_Supplement/Display_items/Main_Figures/Figure5.pdf", plot=eae_change_plot,
       width=11, height=9)
```

```{r}
eae_change$test <- eae_change$`Change in log eae ng`

pair_eae_change <- compare_means(test ~ myaxis, data=eae_change, p.adjust.method = "holm", 
                               method="wilcox.test", ref.group = "placebo\nn=56") %>%
  mutate(y.position=c(7,7.1))


pair_eae_change
eae_change_plot + stat_pvalue_manual(pair_eae_change)
```